### 2015-04-07 pics/front1.tex
front pictures

\begin{tikzpicture}[scale=1.0,>=stealth]
    %\Large
    %\draw [help lines] (0,0) grid (16,15);
% links
    \draw [->,decorate,decoration=snake,diplom2,ultra thick] (0.5,9.2) --
              (2,8.5) node [midway,anchor=south]{$h\nu$};
    \node at (3,8) {\includegraphics[scale=0.10]{nearest_one.png}};
    \draw[->,very thick,diplom3] (4,8.5) -- (5,9.5)
         node [above]{$\color{diplom3}{e^-}$};

    \node at (3,6) {\textbf{Ionization}};

%Mitte
    \draw[->,ultra thick] (4.5,8) -- (6.5,8);
    \node at (8,8) {\includegraphics[scale=0.10]{nearest_two.png}};
    \draw[->,ultra thick,diplom3] (9.5,8) -- (11.5,8)
         node [midway,anchor=south]{$\color{diplom3}{e^-}$};

    \node at (8,6) {\textbf{ICD}};

%unten
% links
    \draw [->,decorate,decoration=snake,diplom2,ultra thick] (0.5,5.2) --
              (2,4.5) node [midway,anchor=south]{$h\nu$};
    \node at (3,4) {\includegraphics[scale=0.110]{next_nearest_one.png}};
    \draw[->,very thick,diplom3] (4,4.5) -- (5,5.5)
         node [above]{$\color{diplom3}{e^-}$};

%Mitte
    \draw[->,ultra thick] (4.5,4) -- (6.5,4);
    \node at (8,4) {\includegraphics[scale=0.11]{next_nearest_two.png}};
    \draw[->,ultra thick,diplom3] (9.5,4) -- (11.5,6.0)
         node [midway,anchor=south,distance=1.5cm]{$\color{diplom3}{e^-}$};

%Rechts
    \node at (14,6) {\includegraphics[scale=0.5,angle=270]{surf3.pdf}};

\end{tikzpicture}