### 2015-05-06 theo_discussion.tex
add more history of structure determination

\section{Discussion}

We have discussed the experimental spectra in iSection \ref{} and
the expected features of spectra for different cluster types in
Section \ref{}. In this section we combine the experimental and theoretical
results in order to determine the structure of the noble gas clusters.