### 2015-07-02 theo_discussion.tex
add beginning of theoretical discussion -> lacks information from experiment

\section{Discussion}

We have discussed the experimental spectra in Section \ref{} and
the expected features of spectra for different cluster types in
Section \ref{}. In this section we combine the experimental and theoretical
results in order to determine the structure of the noble gas clusters.

The simulated secondary electron spectra for the xenon core structures
(see Figure \ref{figure:xe_3_in})
show an ICD peak at higher energies than those structures with added xenon
atoms on the surface. This does not fit to the experimentally determined
spectra and hence we consider an argon core to be more probable for the
small measured clusters.

The experimental spectra in Figure \ref{}
show a broadening of the secondary electron peak with increasing
pressure (?) and hence cluster size just as can be expected from
the theoretically simulated spectra in Figure \ref{figure:surf}.

From the theoretical simulations of larger small clusters
having the same xenon content as experimentally observed a clearly
visible ETMD peak is to be expected as shown in Figure \ref{figure:ar_3_6in}.
However, the experimental spectra do not show a signal in the energy
region of \unit[1.5--6.5]{eV}. From this we conclude that the clusters
measured were smaller than 55 atoms containing only very few xenon atoms.
Additionally, if the cluster contains more than one xenon atom, they are
not expected to be close because of an otherwise visible ETMD peak
(compare Figure \ref{figure:2tops}).

The large clusters with a large xenon content can be expected to have
a xenon core being surrounded by argon layers and hence, both ICD and
ETMD should be visible in the spectrum. Unfortunately, the ICD and ETMD
peaks overlap such that the two processes can not be distiguished.