### 2016-05-23 summary.tex
summary added - version for sending to co-authors

\section{Summary}
%
In summary, we have presented comprehensive experimental and theoretical data for the autoionization of inner valence ionized states in ArXe clusters.
Both ICD and ETMD are allowed for most cases we considered.
Because of energetical reasons ICD requires a separation of the final state vacancies by about 8~\AA or more, which is at least two times the typical Ar-Xe distance.
We found that autoionization has 0.8 to 1.0 efficiency within the experimental accuracy, that is it dominates over other modes of relaxation.
By comparing our measured spectra to calculations, we identified `long range' ICD as the most important decay mode.