### 2016-05-23 abstract.tex

\begin{abstract}
We report about autoionization channels of Ar inner valence ionized states in mixed ArXe clusters. 
The mixed clusters were produced by supersonic coexpansion of the two gases.
Clusters with a range of sizes and compositions were probed. 
The Xe content in the clusters was varied between 10 and 53\,\%. 
Our experimental data obtained by electron, electron coincidence spectroscopy are compared to {\it ab initio} calculations for representative cluster structures.
The combined experimental and theoretical data show that the autoionization of Ar 3s$^{-1}$ in ArXe is dominated by Interatomic Coulombic Decay (ICD) to Xe atoms in the second coordination shell of the originally excited atom.
Electron Transfer Mediated Decay (ETMD) is energetically allowed as well for most cluster structures, but has less intensity in the experimental spectra.
We show experimentally that the relaxation of Ar inner valence states by ICD and ETMD together has an efficiency of unity, within the experimental accuracy, for all clusters expect those with the lowest Xe content.
Decay rates calculated for ETMD favour cluster structures with either a clear seggregation between Ar and Xe fractions, e.g. Xe core, Ar shell systems, or with few Xe atoms singled out in surface sites.
They differ significantly from calculated minimum energy structures for ArXe systems, which might show that the latter structures are not realized in supersonic expansion.
\end{abstract}